---
title: "Model Creation and Testing"
author: "Luis Pedro Insua"
date: "11/13/2020"
output: html_document
---

## Introduction
The objective of this project is to build a model to predict the quality of a dumbbell curl based on data measured by accelerometers.

The data for the project came from the following source:  http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

The direct link to the taining data:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Testing data:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
Loading all packages to be used and set a seed for reproducibility
```{r packages}
if(!require(caret)){ install.packages("caret",repos = "http://cran.us.r-project.org")}
library(caret)
if(!require(randomForest)){ install.packages("randomForest",repos = "http://cran.us.r-project.org")}
library(randomForest)
set.seed(8783)
```

## Loading and processing data
Load data
```{r load and process data, cache=TRUE}
## Load data and delete missing values.
trainURL <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
testURL <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
train <- read.csv(url(trainURL), na.strings=c("NA","#DIV/0!",""))  
test<- read.csv(url(testURL), na.strings=c("NA","#DIV/0!",""))

## Remove variables with no variability.
train <-train[,colSums(is.na(train)) == 0]
test <-test[,colSums(is.na(test)) == 0]

## Remove variables that are not predictors.
train <- train[,-c(1:7)]
test <- test[,-c(1:7)]

## Convert independent variable to factor.
train$classe <- as.factor(train$classe)
```


## Cross validation split             
Split the training data into training and testing subsets.
```{r cross validation split}
intTrain <- createDataPartition(y=train$classe,p=0.7, list=FALSE)
valid.train <- train[intTrain,]
valid.test<- train[-intTrain,]
```

## Prediction Models

### Random Forest Model
Build a random forest model and evaluate its accuracy.
```{r random forest, cache=TRUE}
forest <- randomForest(classe~., data=valid.train, method="class")
pred.forest <- predict(forest, newdata=valid.test)
confusionMatrix(pred.forest,valid.test$classe)
```

### Decision Tree Model
Build a decision tree model and evaluate its accuracy.
```{r decision tree}
tree <- train(classe~., data=valid.train, method="rpart")
pred.tree <- predict(tree, newdata=valid.test)
confusionMatrix(pred.tree,valid.test$classe)
```

### Final Model
Since the Random Forest method was the most accurate in the cross validation, we will use it to create the final model with all the training data.
```{r final model, cache=TRUE}
finalModel <- randomForest(classe~., data=train, method="class")
finalPred <- predict(finalModel, newdata=test)

```

## Works Cited
Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.

Read more: http:/groupware.les.inf.puc-rio.br/har#ixzz4TjqlcsCi
